---
title: "DS Project Team 11"
author: "[Evan Patel, Daksh Verma, Saurabh Rane, Sujay Shah, Tejas Dighe, Maxine Attobrah]"
output: pdf_document
---

# DS Project Team 11
# User input information on their chronic illness from Flaredown app dataset 


```{r setup, include = TRUE, warning = FALSE, message = FALSE}
# libraries needed
library(tidyverse)
library(dplyr)

#data
chronic_illness <- read_csv("chronic_illness_data_split_15050.csv")

```
```{r, warning = FALSE, message = FALSE}

# COLUMNS
# trackable_value: inludes symptom or conditon rating from 0- 4, dose from treatments (units are mg or mcg) [continous variable]
# age
# sex
# country
# checkin_date 
# user_id: idenitfies user?
# trackable_id: idenitfies user?
# condition
# symptom
# weather
# Food : food items that cause conditions?
# treatment
# tag
# HBI

```

## Inital wrangling of daa
The Flaredown app labels users inputs as conditions, symptoms and treatments which was all in the the trackable_type column. In this section we make those labels into columns to get a better understanding of how the conditions, symptoms and treatments affect each other. 


```{r, warning = FALSE, message = FALSE}
# data before
chronic_illness

# dat after making the conditions, symptoms and treatments columns
chronic_illness <- pivot_wider(chronic_illness, names_from=trackable_type, values_from=	
trackable_name)

# print new to csv in my folder
write.csv(chronic_illness, "chronic_illness_data_split_columns_15050_updated.csv")

```



```{r, warning = FALSE, message = FALSE}
# Age

# drop any user that has an unknown age
chronic_illness <- chronic_illness %>%
          drop_na(age)

# display the varying agaes using the app
ggplot(chronic_illness, aes(x=age)) +
  geom_bar()+
  labs(title="Title: Age of people who participated in Flaredown App")

```


```{r, warning = FALSE, message = FALSE}
# country

# drop users that have a unknown location
chronic_illness <- chronic_illness %>%
          drop_na(country)

# display the number of userrs in different countries and which country has the most users
ggplot(chronic_illness, aes(x=country)) +
  geom_bar()+
  labs(title="Title: Country of people who participated in Flaredown App")

```


```{r, warning = FALSE, message = FALSE}

# display what type of people are using the app and who is using the app the most
ggplot(chronic_illness, aes(x=sex, colour = age)) +
  geom_bar()+
  geom_text(aes(label=stat(count)), stat="count", vjust=1.5, colour="red")+
  labs(title="Title: Sex of people who participated in Flaredown App")

```


```{r, warning = FALSE, message = FALSE}

# display ages using the app but color code based on the users sex
ggplot(chronic_illness, aes(x=checkin_date, y=age, colour = sex)) +
  geom_point() +
  labs(title="Title: Age of people who participated in Flaredown App")




# view how which ages report having Ulcerative colitis
chronic_illness %>%
  filter( Condition=="Ulcerative colitis", na.rm=TRUE) %>%
  ggplot(aes(x=checkin_date, y = age)) +
  geom_point()


```
## Display the unique countries, treatments, conditions, symptoms, foods, and tags


```{r, warning = FALSE, message = FALSE}

unique(chronic_illness$country)
```

```{r, warning = FALSE, message = FALSE}
unique(chronic_illness$Treatment)
```

```{r, warning = FALSE, message = FALSE}
unique(chronic_illness$Condition)
```

```{r, warning = FALSE, message = FALSE}
unique(chronic_illness$Symptom)
```

```{r, warning = FALSE, message = FALSE}
unique(chronic_illness$Food)
```

```{r, warning = FALSE, message = FALSE}
unique(chronic_illness$Tag)
```

```{r, warning = FALSE, message = FALSE}

# HBI does not have any value in this data since no one filled it out
unique(chronic_illness$HBI)

```

## RANK showing the most critical items to pay attention to

```{r, warning = FALSE, message = FALSE}

# display the top conditions first
chronic_illness %>%
 group_by(Condition) %>%
                summarize(
                n=n()) %>%
                # arrange so that the biggest 
                # observation is shown first
                arrange(desc(n))


# display most used treatments first
chronic_illness %>%
 group_by(Treatment) %>%
                summarize(
                n=n()) %>%
                # arrange so that the biggest 
                # observation is shown first
                arrange(desc(n))


# dusplay the top symptoms first
chronic_illness %>%
 group_by(Symptom) %>%
                summarize(
                n=n()) %>%
                # arrange so that the biggest 
                # observation is shown first
                arrange(desc(n))


# display which country users are using the app the most
chronic_illness %>%
 group_by(country) %>%
                summarize(
                n=n()) %>%
                # arrange so that the biggest 
                # observation is shown first
                arrange(desc(n))


# display which foods are causing the most reactions
chronic_illness %>%
 group_by(Food) %>%
                summarize(
                n=n()) %>%
                # arrange so that the biggest 
                # observation is shown first
                arrange(desc(n))



# display the top environmental factors that does not occur everyday entered by users 
chronic_illness %>%
 group_by(Tag) %>%
                # move tsi to the front of the dataframe
               # tell me how many observations 
                # there are for each tsi value
                summarize(
                n=n()) %>%
                # arrange so that the biggest 
                # observation is shown first
                arrange(desc(n))

```





